The Types and Driving Forces of Modifications in the Evolution of an API

### 2015-05-18 thesis/paper.tex
Re-wrote rq2 results summary

The results related to \textbf{RQ2} are structured around each of Lehman's laws of software evolution \cite{lehman90sview} and to what extent our findings show that the laws apply to platform APIs. A summary of our findings can be seen in Table \ref{table:lehman}.

@@

              \textbf{Law}                              & \textbf{Applicability}    \\ \midrule
              Continuing Change                         & True                      \\ \hline
              Increasing Complexity                     & Kinda                     \\ \hline
              Self Regulation                           & True                      \\ \hline
              Conservation of Organisational Stability  & False                     \\ \hline
              Conservation of Familiarity               & True                      \\ \hline
              Continuing Growth                         & True                      \\ \hline
              Declining Quality                         & Kinda                     \\ \hline
              Feedback System                           & True                      \\

@@

\smallskip \noindent
\textbf{Increasing Complexity  } \label{sec:law2} 
A majority of the changes introduced in version 2.0 aimed to improve the structure and the way that the API was used. Many of these changes were related to the same design choices, which enforced a certain implementation style on the whole API. One example of this is the change to the interfaces which set a standard for the whole API. During the interview it was also concluded that consistency and following certain standards is an important factor to consider when designing APIs. It was also made clear that a contributing factor for changing certain parts of the API was to bring it into line with the new interface design. The interviewee mentioned that one way of achieving this was to follow guidelines or design best practices, but that the company currently did not make use of these. Even though the importance of maintaining the API in order to reduce complexity was stated, arguments which support that it is of lesser importance than releasing functional content was also presented. One example of where this is evident in the evolution of the API, is in the module that handles the video capturing and video stream. It was decided that this module would not be re-designed to conform with the new interface standard, as the changes to functionality in this module were very minor. Thus, the law of increasing complexity holds for APIs, but there is evidence that suggests that decreasing the complexity is not of the highest priority.

@@

\smallskip \noindent
\textbf{Declining Quality  } 
The interviews established that a groomative motive was not the primary reason for implementing any change. The changes to the example code and the documentation, as well as the re-designed modules are examples of these, where the main motive behind the change was to either prevent unwanted usage or to increase the functional content. These changes did, however, also increase the quality and were a substantial part of the update. A counterargument to this reasoning can be seen in the module which handles the video data and capturing, which was not changed between the versions, as mentioned in Section \ref{sec:law2}. We therefore settle that the law of declining quality can be applied to APIs, but that there is data that implies that reducing quality is not of high importance.

@@

\noindent
\textbf{A majority of the laws apply  } 
Evidence was found for the laws of Continuing Change, Self Regulation, Conservation of Familiarity, Continuing Growth and Feedback System to establish that they apply to APIs. The most prominent explanation for this is that a large part of the introduced changes were related to functionality. It is especially interesting to note that the laws of Self Regulation and Conservation of Familiarity accurately match the API explored in this paper.

\smallskip \noindent
\textbf{Two laws apply to a certain extent  } 
The laws of Increasing Complexity and Declining Quality were found to apply to APIs, but the extent to which they apply might not be as significant for APIs compared to other types of systems. The reason that this doubt exists is because these motives were secondary to the motives of increased functionality. An alternative explanation for these laws being of lesser significance in APIs could be that the general inertia to modify APIs which exists \cite{google_talk, henning2007api, mcdonnell2013empirical, robbes2012developers} might dominate the need to improve simplicity and quality.

\smallskip \noindent
\textbf{One law is not applicable  } 
Strong evidence was found to suggest that the law of Conservation of Organisational Stability does not apply to APIs. Our results not only show that the law is not true for APIs, but that the opposite of what the law states is true for APIs. This is supported by the consensus in literature of the negative effects caused by modifications to APIs \cite{google_talk, henning2007api, mcdonnell2013empirical, robbes2012developers}. The results from our study also revealed that reluctance to change APIs causes the releases of new versions to be few and sporadic.

\smallskip \noindent
\textbf{The importance of a Feedback System  } 
Especially interesting results were found in regards to the law of Feedback System. The importance of having a structured and developed feedback loop was stressed during the interview and it was also expressed as one of the main points of improvement. This suggests that including feedback from more sources and using this feedback to improve the evolution process should be a consideration for APIs.

>>>>

Our analysis of the results related to \textbf{RQ2} showed that seven out of the eight laws \cite{lehman90sview} do apply to APIs. A summary of our findings can be seen in Table \ref{table:lehman}. The evidence found for the laws of continuing change, self regulation, conservation of familiarity, continuing growth and feedback system mainly relate to enhancive changes that add functionality to the API. It was also found that the laws of increasing complexity and declining quality do apply, but that the changes which support these laws usually had a secondary intent related to complexity or quality. The only law which was found not to apply to the API analysed in this study was the law of conservation of organisational stability. Our results rather show the opposite, that the work rate is variant.

In the remainder of this section we will discuss the results related to each law, and discuss if our results show that the laws apply to APIs or not.

@@

              \textbf{Law}                              & \textbf{Applicability}    \\ \midrule
              Continuing Change                         & True                      \\ \hline
              Increasing Complexity                     & True                     \\ \hline
              Self Regulation                           & True                      \\ \hline
              Conservation of Organisational Stability  & False                     \\ \hline
              Conservation of Familiarity               & True                      \\ \hline
              Continuing Growth                         & True                      \\ \hline
              Declining Quality                         & True                     \\ \hline
              Feedback System                           & True                      \\

@@

\smallskip \noindent
\textbf{Increasing Complexity  } \label{sec:law2} 
A majority of the changes introduced in version 2.0 aimed to improve the structure and the way that the API was used. Many of these changes were related to the same design choices, which enforced a certain implementation style on the whole API. One example of this is the change to the interfaces which set a standard for the whole API. During the interview it was also concluded that consistency and following certain standards is an important factor to consider when designing APIs. It was also made clear that a contributing factor for changing certain parts of the API was to bring it into line with the new interface design. The interviewee mentioned that one way of achieving this was to follow guidelines or design best practices, but that the company currently did not make use of these. Although the importance of reducing the complexity was stated, it was made clear during the interview that one module was not updated between the versions, even though it did not follow the new standards of version 2.0. In conclusion, we establish that the law of increasing complexity is true for APIs.

@@

\smallskip \noindent
\textbf{Declining Quality  } 
The interviews established that a groomative motive was not the primary reason for implementing any change. The changes to the example code and the documentation, as well as the re-designed modules are examples of these, where the main motive behind the change was to either prevent unwanted usage or to increase the functional content. These changes did, however, also increase the quality and were a substantial part of the update. Our analysis also shows that changes related to quality are bound to happen, as other types of changes often incorporate quality related aspects. We therefore settle that the law of declining quality applies to APIs.

@@
